First, I would like to thank my advisor, Dr. Hubertus Von Bremen, who guided my thinking and patiently assisted me throughout my graduate program.
I greatly appreciate you pushing me to learn and explore more mathematical ideas during this amazing process.
Because of you, I aim to further my education in applied mathematics as well as computer programming.

I also want to thank my family for supporting me through many sleepless days.
More specifically, I want to thank my mom and dad for helping me stay focused and determined every day.
Furthermore, I am grateful for my fianc\'{e}e, who has sacrificed her time and education for me.
I am excited to return the favor by helping you get your master's degree.

My gratitude goes out to all my close friends and colleagues, Joshua Garcia, Nick Beltz, Ethan Flora, Daniel Silva, Jeffrey Robbins, Seth Ricarte, Jose Contreras, Sara Elakesh, Michael Yates, and Latimer De'Shone Harris-Ward, for all your encouragement and mathematical aid that has made this possible.

I would also like to thank Dr. Berit Givens, Dr. John Rock, and Dr. Jennifer Switkes, who have had the most significant impact on my thinking as a mathematician.
I appreciate your guidance and dedication to helping me understand and apply mathematical concepts.
You all have inspired me to become a better mathematician, and I will always aim to be as knowledgeable as you all.

Lastly, I want to extend my gratitude to Aaron Tiernan, an Area Management Biologist for the ADFG. Thank you for helping me gather information about the salmon population in Bristol Bay.
